WrapITK: Enhanced languages support\\ for the Insight Toolkit

### 2006-06-12 article/Article.tex
add precisions about external projects and ipython prompt and add a full python example
darcs-hash:20060612203858-2fc9d-be9516d2481247d46b76776c724ad684573975b0.gz

Note that ITK already has a wrapping infrastructure, and that WrapITK is based
on it, and use the same tools: CMake \cite{CMakeWebSite}, GCC-XML \cite{GccxmlWebsite}
and CableSwig\footnote{CableSwig is based on a now quite old version of SWIG
\cite{SwigWebSite}.} \cite{CableSwigWebSite}. This project aims to
address the following deficits of the existing wrappers (and others):
\begin{itemize}
  \item  ITK is a huge library, but only a small number of classes are available
in target languages. It become quickly frustrating for the user, especially when
he has to spend lot of time to extend the current set of classes. Even if it is
not yet complete, the WrapITK's set of classes have been highly extended. Moreover,
the user can choose at build time which types and which dimensions he want to wrap.

@@

@@

@@

@@

@@

@@

  \subsection{Contributing to WrapITK}

>>>>

Note that ITK already has a wrapping infrastructure, and that WrapITK is based
on it, and use the same tools: CMake \cite{CMakeWebSite}, GCC-XML \cite{GccxmlWebsite}
and CableSwig\footnote{CableSwig is based on a now quite old version of SWIG
\cite{SwigWebSite}.} \cite{CableSwigWebSite}. This project aims to
address the following deficits of the existing wrappers (and others):
\begin{itemize}
  \item  ITK is a huge library, but only a small number of classes are available
in target languages. It become quickly frustrating for the user, especially when
he has to spend lot of time to extend the current set of classes. Even if it is
not yet complete, the WrapITK's set of classes have been highly extended. Moreover,
the user can choose at build time which types and which dimensions he want to wrap.
With 197 wrapped filters, WrapITK covers 62\% of the available filters.

@@

In this section, we detail python usage. Some of the examples shown here are copied
from the console, and so show the interpreter prompt:

\begin{verbatim}
2> 12+3
2> 15

3> result = 12+3

4> 
\end{verbatim}

In the example above, \verb$12+3$ is what is wrote in the interpreter, and
\verb$15$ is the result. \verb$2>$, \verb$3>$, \verb$4>$ are the prompt
of the interpreter.

@@

     \subsubsection{Full python script examples}

This script is the exact transcription to python of the C++ example 
which can be found at Examples/Filtering/GradientMagnitudeRecursiveGaussianImageFilter.cxx
in the ITK source tree. More information about the filters used can be found in 
the ITK Software Guide \cite{{ITKSoftwareGuide}}, section 6.4.2.

\small \begin{verbatim}
import itk
from sys import argv

InputPixelType = itk.F
OutputPixelType = itk.F

InputImageType = itk.Image[InputPixelType, 2]
OutputImageType = itk.Image[OutputPixelType, 2]

reader = itk.ImageFileReader[InputImageType].New( FileName=argv[1] )

filter = itk.GradientMagnitudeRecursiveGaussianImageFilter[InputImageType, OutputImageType].New(
                     reader,
                     Sigma=float(argv[3]) )
filter.Update();

WritePixelType = itk.UC
WriteImageType = itk.Image[WritePixelType, 2]

rescaler = itk.RescaleIntensityImageFilter[OutputImageType, WriteImageType].New( filter,
                     OutputMinimum=0,
                     OutputMaximum=255 )

writer = itk.ImageFileWriter[WriteImageType].New( rescaler, FileName=argv[2] )

writer.Update();
\end{verbatim} \normalsize

More examples can be found in the directory \verb$Python/Tests$.

@@

External projects are not yet supported for Tcl and Java. See \ref{section:contribute}
to contribute external project support for those languages.

@@

This project is a python only project. It requires to have Numeric, Numarray or numpy installed
on your system, and will let the user convert ITK images to python matrices, and python matrices
to ITK images. It thus provide a bridge between ITK and other great python tools like SciPy
\ref{SciPyWebSite} (and a lot of others).

Once installed, the function are directly available in the \verb$itk$ module - there is nothing
special to import. A \verb$PyBuffer$ template let you choose the type to convert, exactly
like with the ITK classes. You can then use the \verb$GetArrayFromImage()$ and
\verb$GetImageFromArray()$ method to convert respectively a array to an ITK image, and an
ITK image to an array. There is no need to instantiate a \verb$PyBuffer$ object: the methods
are \verb$static$.

\begin{verbatim}
1> import itk

2> reader = itk.ImageFileReader.IUS2.New(FileName='cthead1.png')

3> array = itk.PyBuffer.IUS2.GetArrayFromImage(reader.GetOutput())

4> array
4>
array([[0, 0, 0, ..., 0, 0, 0],
       [0, 0, 0, ..., 0, 0, 0],
       [0, 0, 0, ..., 0, 0, 0],
       ...,
       [0, 0, 0, ..., 0, 0, 0],
       [0, 0, 0, ..., 0, 0, 0],
       [0, 0, 0, ..., 0, 0, 0]], type=UInt16)

5> image = itk.PyBuffer.IUS2.GetImageFromArray(array)

6> image
6> <C itk::SmartPointer<(itk::Image<(unsigned short,2)>)> instance at _50684208_p_itk__SmartPointerTitk__ImageTunsigned_short_2_t_t>
\end{verbatim}

Because PyBuffer is a python only external project, its directory structure is very
simple - there is no subdirectory. This external project should be used as an example
for all the languages specific external projects.

@@

This project provide new feature for all the languages, including C++.
Its directory structure reflect that.

\begin{verbatim}
|-- CMakeLists.txt
|-- Wrapping
|   |-- CMakeLists.txt
|   |-- Python
|   |   |-- CMakeLists.txt
|   |   |-- Tests
|   |   |   |-- CMakeLists.txt
|   |   |   |-- CannyEdgeDetectionImageFilter.py
|   |   |   `-- simpleItkVtkPipeline.py
|   |   `-- itkvtk.py
|   |-- itkvtk.swg
|   |-- wrap_itkImageToVTKImageFilter.cmake
|   `-- wrap_itkVTKImageToImageFilter.cmake
|-- images
|   `-- cthead1.png
`-- src
    |-- itkImageToVTKImageFilter.h
    |-- itkImageToVTKImageFilter.txx
    |-- itkVTKImageToImageFilter.h
    `-- itkVTKImageToImageFilter.txx
\end{verbatim}

The C++ source files are in
directory src, while the files needed for WrapITK are in Wrapping.
The CMakeLists.txt file in the root of the project includes the
Wrapping sub directory only if the user ask for it with the option
BUILD\_WRAPPERS. Some python specific code can be found in Wrapping/Python
directory, and some python tests in Wrapping/Python/Tests.
The itkvtk.swg file contains the typemaps required to return vtk
objects.
The images directory contains the images used for the tests. Putting
the images in this directory rather than in the root of the project
prevent overriding the reference files during the test, if the build
is done in the source tree.
The project should also provide C++ tests - it is not done yet.

@@

  \subsection{Contributing to WrapITK}
\label{section:contribute}
